---
title: "part 2 chi square with lizards' tails"
author: "Iris Foxfoot"
date: "12/2/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)
library(broom)
```

## Read in data

```{r}
lizards <- read.csv(here("data", "lizards.csv"))
```

Get counts of lizards by species
```{r}
lizards %>% 
  count(spp) %>% 
  arrange( -n)
```

## Make a subset

- UTST Lizards

-Sites: IPBE (enclosure grassland), RABB (rabbits mesquite site), "EAST" (east tarbush site)

-Tail condition is recorded (W = whole, B = Broken)

```{r}
utst <- lizards %>% 
  filter(spp == "UTST") %>% 
  filter(site %in% c("IBPE", "RABB", "EAST")) %>% 
  filter(tail %in% c("B", "W"))
```

## Find counts and proportions of lizards with broken or whole tails at each site

```{r}
utst %>% 
  count(site, tail)
```

```{r}
tail_counts <- utst %>% 
  janitor::tabyl(site, tail)

tail_proportions <- tail_counts %>% 
  janitor::adorn_percentages() %>% 
  janitor::adorn_pct_formatting(digits = 2) %>% 
janitor::adorn_ns()
```

## Convert to a contingency table

use `column_to_rowname()`

```{r}
lizard_ct <- tail_counts %>% 
  column_to_rownames(var = "site")
```

## Chi Square

```{r}
lizard_x2 <- chisq.test(lizard_ct)
lizard_x2

lizard_tidy <- broom::tidy(lizard_x2)
```

Using a sig of 0.05, yes we have enough evidence to reject the null hypothesis that these proportions are from the same group. There is a significant association between site and tail breakage. Site and lizard tail breakage are not independent.

#In Line referencing
there is a significant assoication (ie non-independence) btween site and lizard tail breakage ($\chi$^2^(`r lizard_tidy$parameter`) = `r lizard_tidy$statistic`, *p* = `r lizard_tidy$p.value`).
